On discretely entropy conservative and entropy stable discontinuous Galerkin methods

### 2018-01-17 docs/splitform/main.tex
updating bbwadg cfd files

\begin{abstract}
High order methods based on diagonal-norm summation by parts operators can be shown to satisfy a discrete conservation or dissipation of entropy for nonlinear systems of hyperbolic PDEs \cite{fisher2013high, carpenter2014entropy}.  These methods can also be interpreted as nodal discontinuous Galerkin methods with diagonal mass matrices \cite{gassner2016split, gassner2016well, wintermeyer2017entropy, chen2017entropy}.  In this work, we describe how to construct discretely entropy conservative schemes to a more general class of DG methods using flux differencing, quadrature-based projections, and specific DG differentiation operators.  This approach also recovers existing methods for Burgers' equation involving dense norm and generalized SBP operators without boundary nodes \cite{ranocha2016summation, ranocha2017extended}.  Numerical experiments confirm the stability and high order accuracy of the proposed methods for the one-dimensional compressible Euler equations.  
\end{abstract}
\end{frontmatter}

@@

The extension of entropy conservative schemes to unstructured high order methods was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can also be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.  The proof of entropy conservation relies on the presence of a diagonal mass matrix, the summation by parts (SBP) property \cite{gassner2013skew}, and  a concept referred to as flux differencing.  Similar entropy-stable schemes have been constructed for the shallow water and MHD equations \cite{gassner2016well,  wintermeyer2017entropy, winters2017uniquely}.  More recently, high order entropy conservative and entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.

@@

This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.  In order to generalize beyond diagonal norm SBP-DG methods, we will consider non-lumped (dense) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space), which are also related to dense norm and generalized SBP operators \cite{ranocha2016summation, ranocha2017extended, ranocha2017comparison}.  The work presented in this manuscript is intended to serve as a stepping stone to the construction of entropy stable DG methods on general elements in two and three dimensions.  We will present proofs of discrete entropy stability in terms of integrals and $L^2$ projections, which we assume are computed using an appropriate quadrature rule.  We stress that, while this work adopts a continuous notation involving integrals rather than discrete sums, the proofs rely only on properties of quadrature-based integration and quadrature-based $L^2$ projection.

>>>>

\begin{abstract}
High order methods based on diagonal-norm summation by parts operators can be shown to satisfy a discrete conservation or dissipation of entropy for nonlinear systems of hyperbolic PDEs \cite{fisher2013high, carpenter2014entropy}.  These methods can also be interpreted as nodal discontinuous Galerkin methods with diagonal mass matrices \cite{gassner2016split, gassner2016well, wintermeyer2017entropy, chen2017entropy}.  In this work, we describe how to construct discretely entropy conservative schemes to a more general class of DG methods using flux differencing, quadrature-based projections, and specific DG differentiation operators.  This approach also recovers existing methods for Burgers' equation involving dense norm and generalized SBP operators without boundary nodes \cite{fernandez2014generalized, ranocha2016summation, ranocha2017extended}.  Numerical experiments confirm the stability and high order accuracy of the proposed methods for the one-dimensional compressible Euler equations.  
\end{abstract}
\end{frontmatter}

@@

The extension of entropy conservative schemes to unstructured high order methods was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can also be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.  The proof of entropy conservation relies on the presence of a diagonal mass matrix, the summation by parts (SBP) property \cite{gassner2013skew}, and  a concept referred to as flux differencing.  Similar entropy-stable schemes have been constructed for the shallow water and MHD equations \cite{gassner2016well,  wintermeyer2017entropy, winters2017uniquely}.  More recently, high order entropy conservative and entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional, crean2017high}.

@@

This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.  In order to generalize beyond diagonal norm SBP-DG methods, we will consider non-lumped (dense) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space), which are also related to dense norm and generalized SBP operators \cite{fernandez2014generalized, ranocha2016summation, ranocha2017extended, ranocha2017comparison}.  The work presented in this manuscript is intended to serve as a stepping stone to the construction of entropy stable DG methods on general elements in two and three dimensions.  We will present proofs of discrete entropy stability in terms of integrals and $L^2$ projections, which we assume are computed using an appropriate quadrature rule.  We stress that, while this work adopts a continuous notation involving integrals rather than discrete sums, the proofs rely only on properties of quadrature-based integration and quadrature-based $L^2$ projection.
